More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-07-24 2016-Weak-FH-Bisim/WeakBisim.tex
reorga OT  / pnets et un peu de texte

% !TEX TS-program = pdflatex

@@

\paragraph{Contribution}

@@

\TODO{informal semantics of pnets}

@@

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $\beta_j$, $s$, $\alpha$. \Post\ is a set of 
	equations 
	that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $({x_k\gets e_k})^{k\in K}$ 
	where $x_k$ are variables of $s'$, and $e_k$ are expressions over the other 
	variables of the open transition. Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}

\begin{definition}[Open automaton]
	\label{def:open-automaton}
	An \emph{open automaton} is a structure\\ $A =
	<J,\mathcal{S},s_0,\mathcal{T}>$ where:
	\begin{itemize}
		\item[$\bullet$]   $J$ is a  set of indices,
		\item[$\bullet$]   $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
		\item[$\bullet$] $\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist  $J'$ with  $J'
		\subseteq J$, such that $t$ is an open transition over  $J'$
		and  $\mathcal{S}$.

@@

\subsection{Computing and using open automata}
In this section we present a simple algorithm to construct the open
automaton representing the behaviour of an open pNet, and we prove that
under reasonable conditions this automaton is finite.

@@

\section{Bisimulation}
\label{section:bisimulation}

	Now we use our symbolic operational semantics to define a notion of
	strong (symbolic) bisimulation. Moreover this equivalence is decidable whenever we 
	have some
	decision procedure on the predicates of the action algebra.

%% , that we prove compatible with the
%% composition of open pNets. \TODO{do we really know this ?}

The equivalence we need is a strong bisimulation between
pNets having exactly the same Holes with the same sorts, but using a
flexible matching 
between open transition, to accommodate comparisons between pNet
expressions with different architectures.
 We name it FH-bisimulation,
 as a short cut for the ``Formal Hypotheses'' manipulated in the
 transitions, but also as a reference to the work of De Simone~\cite{deSimone85},
 that pioneered this idea.

%Where $\symb{Subst}(\Pred,\Post)$ is the parallel substitution of all
%assigned variables.
Two pNets are FH-bisimilar if there exist a relation between their associated 
automata that is an FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is true.
% \TODO{do we want formulas for this?}
 \end{definition}
Classically, $\Pred_{s',t_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}$
applies in parallel the  
substitutions $\Post_{OT}$ and $\Post_{OT_x}$ (parallelism is crucial
inside each $\Post$ set but not between  $\Post_{OT}$ and
$\Post_{OT_x}$ that are independent), applying the assignments of the involved rules.

\medskip
If we assume that everything is finite (states and transitions in the
open automata, and the predicates in $\mathcal{R}$, then it is easy to
prove that it is decidable whether a relation is a 
FH-bisimulation, provided the logic of the predicates is decidable (see \cite{henrio:Forte2016}).

\section{Composability}

The main interest of our symbolic approach is to define a method to
prove properties directly on open structures, that will be preserved
by any correct instantiation of the holes. In this section we define a
composition operator for open pNets, and we prove that it preserves
FH-bisimulation. More precisely, one can define two preservation
properties, namely 1) when one hole of a pNet is filled by two bisimilar other (open) pNets; and 2) when the same hole in two bisimilar pNets are
filled by the same pNet, in other words, composing a pNet with two
bisimilar contexts. The general case will be obtained by
transitivity of the bisimulation relation. We concentrate here on the
second property, that is the most interesting.

>>>>

% !TEX TS-program = pdflatex
% !TeX spellcheck = en_GB

@@

\paragraph{Contribution}
The main interest of our symbolic approach is to define a method to
prove properties directly on open structures, that will be preserved
by any correct instantiation of the holes.

@@

The informal semantics of \pNet s is as follows. \pLTS s behave more or less like  classical automata with conditional branching and variables. The actions on the \pLTS s can send or receive values, potentially modifying the value of variables. 
\pNet s are synchronisation entities: a \pNet composes several subnets and  synchronisation vectors define how the subnets interact, where a subnet is either a \pNet or a \pLTS. Synchronisation between subnets is defined by synchronisation vectors that express how an action of a subnet can be synchronised with actions of other subnet, and how the resulting synchronised action is visible to the outside of the \pNet. Synchronisation vectors also allow to explicitly export an action of a subnet to the next level, or to allow to subnets to synchronise and only exporting a non0observable action (provided such an non-visible action labels exist in the action algebra. Finally, a \pNet can leave subnets undefined and instead declare holes with a well-defined signature. Holes can then be filled with a subnet. This is defined as follows.

@@

\subsection{Open Automata: A model of process composition}
\label{sec:OT}

The semantics of open \pNet s will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications. This section defines open automata and a bisimulation theory for them. Open automata are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on variables (potentially with only symbolic values).
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $\beta_j$, $s$, $\alpha$. \Post\ is a set of 
	equations 
	that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $({x_k\gets e_k})^{k\in K}$ 
	where $x_k$ are variables of $s'$, and $e_k$ are expressions over the other 
	variables of the open transition. Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}

An open automaton is then simply an automaton where each transition is an open transition.
\begin{definition}[Open automaton]
	\label{def:open-automaton}
	An \emph{open automaton} is a structure\\ $A =
	<J,\mathcal{S},s_0,\mathcal{T}>$ where:
	\begin{itemize}
		\item[$\bullet$]   $J$ is a  set of indices,
		\item[$\bullet$]   $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
		\item[$\bullet$] $\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist  $J'$ with  $J'
		\subseteq J$, such that $t$ is an open transition over  $J'$
		and  $\mathcal{S}$.

\subsection{Bisimulation for open Automata}
\label{section:bisimulation}

The equivalence we need is a strong bisimulation between
pNets having exactly the same Holes with the same sorts, but using a
flexible matching 
between open transition, to accommodate comparisons between pNet
expressions with different architectures.

We define now a bisimulaiton relation adapted to open automata and their parametric nature. The relation relates states of the open automaton and states equivalence between the open transitions between the states. Its key characteristics are 1) the introduction of predicates in the bisimulation relation: as states may contain variables, relation between states may depend on the value of the variables; 2) the bisimulation property relates elements of the open transitions and take into account predicates over variables, actions of the holes, and state modifications.
 We name it FH-bisimulation,
 as a short cut for the ``Formal Hypotheses'' manipulated in the
 transitions, but also as a reference to the work of De Simone~\cite{deSimone85},
 that pioneered this idea.

%Where $\symb{Subst}(\Pred,\Post)$ is the parallel substitution of all
%assigned variables.

% \TODO{do we want formulas for this?}
 \end{definition}
Classically, $\Pred_{s',t_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}$
applies in parallel the  
substitutions $\Post_{OT}$ and $\Post_{OT_x}$ (parallelism is crucial
inside each $\Post$ set but not between  $\Post_{OT}$ and
$\Post_{OT_x}$ that are independent), applying the assignments of the involved rules.
We can prove that such a bisimulation si an equivalence relation:

If we assume that everything is finite (states and transitions in the
open automata, and the predicates in $\mathcal{R}$, then it is easy to
prove that it is decidable whether a relation is a 
FH-bisimulation, provided the logic of the predicates is decidable (see \cite{henrio:Forte2016}).

@@

@@

\subsection{Computing and using open automata}
\TODO{REMOVE?}
In this section we present a simple algorithm to construct the open
automaton representing the behaviour of an open pNet, and we prove that
under reasonable conditions this automaton is finite.

@@

\subsection{Bisimulation for open \pNet s and Composability}
\label{section:bisimulation-PN}
As  our symbolic operational semantics provides an open automaton, we can apply the notion of
	strong (symbolic) bisimulation on automata to open \pNet s:
\begin{definition}[FH-bisimulation for open \pNet s]\label{def:bisim-pnets}
Two pNets are FH-bisimilar if there exist a relation between their associated 
automata that is an FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is true.
\end{definition}
We can now prove that \pNet\ composition  preserves
FH-bisimulation. More precisely, one can define two preservation
properties, namely 1) when one hole of a \pNet\ is filled by two bisimilar other (open) pNets; and 2) when the same hole in two bisimilar pNets are
filled by the same pNet, in other words, composing a pNet with two
bisimilar contexts. The general case will be obtained by
transitivity of the bisimulation relation.
